---
title: Yampolsky et al. (2014) Adaptive phenotypic plasticity and local adaptation
  for temperature tolerance in freshwater zooplankton
author: "Tad Dallas"
output:
  html_document:
    toc: yes
  pdf_document:
    fig_caption: yes
    toc: yes
---


\newpage 

# Introduction 

```{r eval=TRUE, echo=TRUE}
 # find all csv files in data directory, and read files in as list elements
 data <- lapply(paste('../data/', dir('../data', '\\.csv'), sep=''), read.csv, 
                stringsAsFactors=FALSE)
 # make colnames consistent
 colnames(data[[2]])[1] <- 'Clone'

 names(data) <- c('HbPublic', 'SuppTable1', 'DmagnaPublic')
```



# Data preprocessing

Clone names are not conserved across datasets. Also, `Clone` and `CloneID` are two columns that pertain to the same information. Also, clone names don't sync up, which may be the reason for the lack of reproducibility for Tables 2 and 4.

```{r eval=TRUE, echo=TRUE}

data[[3]][which(data[[3]][,'Clone'] == 'Xinb3'),'Clone'] <- 'Xinb-3'
data[[3]][which(data[[3]][,'Clone'] == 'Iinb1'),'Clone'] <- 'Iinb-1'



```

\newpage

# Figure 1


```{r eval=TRUE, echo=TRUE, fig.width=5, fig.cap="Time until immobilization as a function of the warmest month average high."}

getFig1 <- function(data, useExisting = TRUE){
  
  if(useExisting == FALSE){
   require(dplyr)
   hotData <- data[[3]][which(data[[3]][['AccT']] == 28),]
   coldData <- data[[3]][which(data[[3]][['AccT']] == 20),]
   hotRet <- as.data.frame(hotData %>%
                          group_by(Clone) %>%
                          summarize(mn = mean(logTimm), 
                                    se = sd(logTimm)/sqrt(length(logTimm))))
   coldRet <- as.data.frame(coldData %>%
                            group_by(Clone) %>%
                            summarize(mn = mean(logTimm), 
                                      se = sd(logTimm)/sqrt(length(logTimm))))
   aht <- c(23, 25, 20,  NA, NA, 21, 21, 27, 23, 21, 24, 31, 30, 30, 
          24, 20, 17, 17, 23, 25, 21, 25, 20) 
 
   names(aht) <- c("BE-OM-2", "CH-H-149", "EL-75-69", "ELNA-1", "F1", "FAT-1-6", 
                   "FHS-2-11-8", "FRC-1-1", "FSP-1-16-2", "FUT-1-2-1", "Iinb-1", 
                   "IR-1-7", "ISR-1-8", "Israel-8", "Mu-11", "N-47-20", "RBol-1-1", 
                   "RKOR-1-1", "RMZ-009", "RYAK", "SE-G1-9", "WR-G1-9", "Xinb-3")
   plot(aht, coldRet[,2], pch=21, las=1, ylim=c(3.5, 5.5), tck=0.01, bg = 'white',
        ylab= expression(paste('log(T'[imm],')')), 
        xlab=expression(paste('warmest month average high, T (', ~degree ,'C)')))
  
   segments(x0 = aht, y0 = coldRet[,2] + coldRet[,3], y1 = coldRet[,2] - coldRet[,3])
   abline(lm(coldRet[,2] ~ aht), lwd=2, lty=2 )
   points(aht, coldRet[,2], pch = 21, bg='white')
  
   points(aht, hotRet[,2], pch = 16)
   segments(x0 = aht, y0 = hotRet[,2] + hotRet[,3], y1 = hotRet[,2] - hotRet[,3])
   abline(lm(hotRet[,2] ~ aht), lwd=2)
  }
  
  if(useExisting){
   aht <- data[[2]][,6]
   logTimm <- data[[2]][,16]
   plot(aht, logTimm, pch=21, las=1, ylim=c(3.5, 5.5), tck=0.01, bg = 'white',
        ylab= expression(paste('log(T'[imm],')')), 
        xlab=expression(paste('warmest month average high, T (', ~degree ,'C)')))
  
   segments(x0 = aht, y0 = logTimm + data[[2]][,18], y1 = logTimm - data[[2]][,18])
   abline(lm(logTimm ~ aht), lwd=2, lty=2 )
   points(aht, logTimm, pch = 21, bg='white')
  
   logTimmH <- data[[2]][,17]
   points(aht, logTimmH, pch = 16)
   segments(x0 = aht, y0 = logTimmH + data[[2]][,19], y1 = logTimmH - data[[2]][,19])
   abline(lm(logTimmH ~ aht), lwd=2)}
}

# this is using the data from the supplied data[[2]] data.frame
getFig1(data)

# this is attempting to calculate means from the data in data[[3]]. I'm missing something.
getFig1(data, useExisting = FALSE)

# excluding data not from the "Main" experiment makes the results look more similar. 
test <- data
test[[3]] <- test[[3]][-which(test[[3]]$Experiment != 'Main'),]
getFig1(test, useExisting=FALSE)
```




\newpage

# Table 2

```{r eval=TRUE, echo=TRUE, results='asis', message=FALSE}
library(xtable)
glmTimm <- aov(lm(c(data[[2]][,16], data[[2]][,17]) ~ rep(data[[2]][,6],2) *
                    c(rep(20,nrow(data[[2]])), rep(28, nrow(data[[2]])))))
glmTimmTable <- matrix(unlist(summary(glmTimm)), ncol=5)

rownames(glmTimmTable) <- c("AHT_warmest", "AccT", "Acct * AHT_warmest", "Residuals")
colnames(glmTimmTable) <- c('df', 'SS', 'MS', 'F', 'p-value')

print(xtable(glmTimmTable), floating.placement='!', type='html')
```



\newpage

#  Figure 2

I took a few liberties and added some pretty colors instead of the cross hatch black and white from the original paper. 

```{r eval=TRUE, echo=TRUE, message=FALSE, comment='', fig.cap='Time until immobilization as a function of acclimization temperature (20 or 28 C) under both high (darker bars) and limited (lighter bars) food resource levels for both short-term (8h; panel _a_) and long-term (1 generation; panel _b_) nutritional manipulation treatments.'}
getFig2 <- function(data, color = rainbow(4),  ...){
 require(dplyr)
 short <- data[[3]][which(data[[3]][,'Experiment'] == 'Short-term feeding'),]
 long  <- data[[3]][which(data[[3]][,'Experiment'] == 'Long-term feeding'),]
  
 shBarData <- short %>% 
   group_by(Treatment, AccT) %>% 
   summarize(mean(logTimm), sd(logTimm)/sqrt(length(logTimm)))
 
 lgBarData <- long %>% 
   group_by(Treatment, AccT) %>% 
   summarize(mean(logTimm), sd(logTimm)/sqrt(length(logTimm)))
  
 layout(matrix(c(1,2),ncol=2))
 shBar <- barplot(shBarData[[3]][c(3,1,4,2)], space = c(0.2, 0, 1, 0), 
                  las=1, tck=0.01, col=color, ylab=expression(paste('log(T'[imm],')')),
                  xlab=expression(paste('acclimitization temperature (', ~degree ,'C)')),
                  ...)
 legend('topright', expression(italic('a')), bty='n', cex=1.25)
 axis(1, labels=c('20', '28'), at = c(1.2, 4.6), tck=0.01)
 
 segments(x0 = shBar, 
          y0 = shBarData[[3]][c(3,1,4,2)] + shBarData[[4]][c(3,1,4,2)],
          y1 = shBarData[[3]][c(3,1,4,2)] - shBarData[[4]][c(3,1,4,2)])
 
 lgBar <- barplot(lgBarData[[3]][c(3,1,4,2)], space = c(0.2, 0, 1, 0), 
                  las=1, tck=0.01, col=color, ylab='', 
                  xlab=expression(paste('acclimitization temperature (', ~degree ,'C)')),
                  ...)
 legend('topright', expression(italic('b')), bty='n', cex=1.25)
 axis(1, labels=c('20', '28'), at = c(1.2, 4.6), tck=0.01)

 segments(x0 = lgBar, 
          y0 = lgBarData[[3]][c(3,1,4,2)] + lgBarData[[4]][c(3,1,4,2)],
          y1 = lgBarData[[3]][c(3,1,4,2)] - lgBarData[[4]][c(3,1,4,2)])
 
}

prettyColors <- c('#dbed9f', '#a5cd26', '#ed9fdb', '#cd26a5')
getFig2(data, color = prettyColors, xpd=FALSE, ylim=c(3,5.5))
```



\newpage

# Table 3


```{r eval=TRUE, echo=TRUE, results='asis', message=FALSE}
getTable3 <- function(data, type='html'){
  require(xtable)
  datS <- data[[3]][which(data[[3]]$Experiment == 'Short-term feeding'),]
  datL <- data[[3]][which(data[[3]]$Experiment == 'Long-term feeding'),]
  
  glmS <- aov(glm(datS$logTimm ~ datS$AccT * datS$Treatment))
  glmS2 <- matrix(unlist(summary(glmS)), ncol=5)
  rownames(glmS2)<- c("AccT", "food", "Acct * food", "Residuals")
  colnames(glmS2)<- c('df', 'SS', 'MS', 'F', 'p-value')
  
  glmL <- aov(glm(datL$logTimm ~ datL$AccT * datL$Treatment))
  glmL2 <- matrix(unlist(summary(glmL)), ncol=5)
  rownames(glmL2)<- c("AccT ", "food ", "Acct * food ", "Residuals ")
  colnames(glmL2)<- c('df', 'SS', 'MS', 'F', 'p-value')
  
  glmTable <- rbind(rep(NA, 5), glmS2, rep(NA, 5), glmL2)
  
  rownames(glmTable)[c(1,6)] <- c("short-term food manipulation treatment", 
                                  "long-term food manipulation treatment")
  
  print(xtable(glmTable), floating.placement='!', 
        comment = FALSE, digits = 2, type=type)
  }

getTable3(data, type='html')
```

\newpage

# Figure 3

Still not quite there. Need the `aht` values that sync up nicely with their respective clone. 

```{r eval=TRUE, echo=TRUE, fig.height=9, fig.width=3}
getFig3 <- function(data, aht, plotIt=TRUE, ...){
  require(dplyr)
  hotData <- data[[3]][which(data[[3]][['AccT']] == 28),]
  coldData <- data[[3]][which(data[[3]][['AccT']] == 20),]
  retHot <- as.data.frame(hotData %>%
                          group_by(Clone) %>%
                          summarize(mn = mean(logTimm), 
                                    se = sd(logTimm)/sqrt(length(logTimm))))
  
  retCold <- as.data.frame(coldData %>%
                           group_by(Clone) %>%
                           summarize(mn = mean(logTimm), 
                                    se = sd(logTimm)/sqrt(length(logTimm))))
  hotData1 <- data[[1]][which(data[[1]][['AccT']] == 28),]
  coldData1 <- data[[1]][which(data[[1]][['AccT']] == 20),]
    
  h576Hot <-  as.data.frame(hotData1 %>%
                            group_by(Clone) %>%
                            summarize(mn = mean(dA576normTProt,na.rm=TRUE), 
                                     se = sd(dA576normTProt,na.rm=TRUE)
                                    /sqrt(length(na.omit(dA576normTProt)))))
    
  h576Cold <-  as.data.frame(coldData1 %>%
                             group_by(Clone) %>%
                             summarize(mn = mean(dA576normTProt,na.rm=TRUE), 
                                      se = sd(dA576normTProt,na.rm=TRUE)
                                      /sqrt(length(na.omit(dA576normTProt)))))
  
   #lifted from manuscript text (where is this in the data?)
   aht <- c(23, 25, 20,  NA, NA, 21, 21, 27, 23, 21, 24, 31, 30, 30, 
          24, 20, 17, 17, 23, 25, 21, 25, 20) 
 
   names(aht) <- c("BE-OM-2", "CH-H-149", "EL-75-69", "ELNA-1", "F1", "FAT-1-6", 
                   "FHS-2-11-8", "FRC-1-1", "FSP-1-16-2", "FUT-1-2-1", "Iinb-1", 
                   "IR-1-7", "ISR-1-8", "Israel-8", "Mu-11", "N-47-20", "RBol-1-1", 
                   "RKOR-1-1", "RMZ-009", "RYAK", "SE-G1-9", "WR-G1-9", "Xinb-3")
  if(plotIt == TRUE){
    #Time to make the mega-plot
    layout(matrix(c(1,2,3), ncol=1))
    par(mar=c(6,4,0,0))
    #panel a
    plot(h576Hot[,2], retHot[,2], pch=21, las=1, xlim=c(0,0.1), ylim=c(3.5, 5.5), 
         tck=0.01, bg = 'white',
         ylab= expression(paste('log(T'[imm],')')), 
         xlab= expression(paste('H'[576])))
    segments(x0 = h576Hot[,2], y0 = retHot[,2]+retHot[,3], y1 = retHot[,2] - retHot[,3])
    segments(x0 = h576Hot[,2]+h576Hot[,3], y0 = retHot[,2], x1 = h576Hot[,2]-h576Hot[,3])
    points(h576Hot[,2], retHot[,2], pch = 21, bg='white')
  
    points(h576Cold[,2], retCold[,2], pch = 16)
    segments(x0 = h576Cold[,2], y0 = retCold[,2]+retCold[,3], 
             y1 = retCold[,2] - retCold[,3])
  
    segments(x0 = h576Cold[,2] + h576Cold[,3], y0 = retCold[,2], 
             x1 = h576Cold[,2]-h576Cold[,3])
    legend('topright', expression(italic('a')), bty='n', cex=1.25)
  
    #panel b
    plot(aht, h576Hot[,2] - h576Cold[,2], pch=21, las=1, tck=0.01, ylim=c(-0.01, 0.1), 
         bg = 'white', ylab= expression(paste(Delta, 'H'[576])), 
         xlab=expression(paste('AHT_warmest (', ~degree ,'C)')), ...)
    abline(lm((h576Hot[,2] - h576Cold[,2]) ~ aht)) 
    legend('topright', expression(italic('b')), bty='n', cex=1.25)
  
    #panel c
    plot(aht, h576Hot[,2], pch=21, las=1, tck=0.01, bg = 'white', ylim=c(-0.01, 0.1),
         ylab = expression(paste('H'[576])), 
         xlab = expression(paste('AHT_warmest (', ~degree ,'C)')), ...)
    segments(x0 = aht, y0 = h576Hot[,2] + h576Hot[,3], 
             y1 = h576Hot[,2] - h576Hot[,3])
    points(aht, h576Hot[,2], pch = 21, bg='white')
  
    points(aht, h576Cold[,2], pch = 16)
    segments(x0 = aht, y0 = h576Cold[,2] + h576Cold[,3], 
             y1 = h576Cold[,2] - h576Cold[,3])
    legend('topright', expression(italic('c')), bty='n', cex=1.25)
  }  
 
  if(plotIt == FALSE){
    ret  <- rbind(retHot, retCold) 
    h576 <- rbind(h576Hot, h576Cold)
    acct <- c(rep(28, nrow(retHot)), rep(20,nrow(retCold)))
    return(list(Timm=ret, h576=h576, AccT=acct, aht = aht))
  }
}

getFig3(data, aht)

```


\newpage

# Table 4

Wrong. 

```{r eval=TRUE, echo=TRUE, results='asis', message=FALSE}
getTable4 <- function(data, type='html'){
  require(xtable)
  raw <- getFig3(data, aht, plotIt = FALSE)
  glmH576 <- aov(glm(log(raw[['h576']][,2]) ~ rep(raw[['aht']],2) * raw[['AccT']]))
  glmH576Table <- matrix(unlist(summary(glmH576)), ncol=5)
  rownames(glmH576Table) <- c("AHT_warmest", "AccT", "Acct * AHT_warmest", "Residuals")
  colnames(glmH576Table) <- c('df', 'SS', 'MS', 'F', 'p-value')
  print(xtable(glmH576Table), floating.placement='!', comment=FALSE, type=type)  
}

getTable4(data, type='html')
```
































```{r eval=FALSE, echo=FALSE}
# First attempt at just using Dryad data to fix non-matching columns. 
# Got frustrated at this silliness

sort(unique(data[[2]][,'Clone']))
sort(unique(data[[3]][,'Clone']))

# These are poor inferences at Clone names.

fixedNames <- matrix(c('BE-OM-2', 'BE-OM-2',
'BY-G1-9', 'SE-G1-9',
'CH-H-149', 'CH-H-149',
'DE-Iinb-1', 'Iinb-1',
'DE-K35-Mu11', 'Mu-11',
'FI-FAT-1-6', 'FAT-1-6',
'FI-FHS2-11-8', 'FHS-2-11-8',
'FI-FSP1-16-2', 'FSP-1-16-2',
'FI-FUT1-2-1', 'FUT-1-2-1',
'FI-N-47-20',  'N-47-20',
'FI-Xinb-3', 'Xinb-3',
'FR-C1-1', 'F1',
'GB-EL75-69', 'EL-75-69',
'IL-M1-8',  'Israel-8',
'IR-GG1-7', 'IR-1-7',
'IT-ISR1-8', 'ISR-1-8',
'KE-1-1', 'FRC-1-1',
'RU-BOL1-1', 'RBol-1-1',
'RU-KOR1-1', 'RKOR-1-1',
'RU-RM1-009', 'RMZ-009',
'RU-YAK1-1', 'RYAK',
'SE-G1-9', 'SE-G1-9'), ncol=2, byrow=TRUE)

for(i in 1:nrow(fixedNames)){
  data[[2]]$Clone[which(data[[2]]$Clone == fixedNames[i,1])] <- fixedNames[i,2]
}



```



```{r eval=FALSE, echo=FALSE, results='asis', message=FALSE}
library(xtable)
glmTimm <- getFig1(data, aht, plotIt=FALSE)
glmTimmTable <- as.data.frame(summary(glmTimm)$coefficients)[-1,]
rownames(glmTimmTable)<- c("AHT_warmest", "AccT", "Acct * AHT_warmest")
print(xtable(glmTimmTable), floating.placement='!', type='html')
```





